Depth and thickness of tectonic tremor in the northeastern Olympic Peninsula

### 2021-06-20 paper/agujournaltemplate.tex
Last changes to paper

For each array, and every 5 km by 5 km grid cell located within 25 km, we analyze the 1-minute long time windows corresponding to all the tremor epicenters located within the grid cell. For each three-component seismic station and each channel, we detrended the data, tapered the first and last 5 seconds with a Hann window, removed the instrument response, bandpass filtered between 2 and 8 Hz, and resampled the data to 20 Hz. All these preprocessing operations were done with the Python package obspy. For each seismic station and each one-minute-long time window, we cross correlated the vertical component with the East-West component and with the North-South component. Then, we stacked the cross correlation functions over all the seismic stations of the array. We call this stack $S_{i j} (\tau)$ for time lag $\tau$, the $i$-th channel ($i$=1 or 2 for the North or East channel) and the $j$-th one-minute time window for a given epicenter grid/array pair. At each step of this method we experimented with a linear stack, a $n$th-root stack, and a phase-weighted stack ~\cite{SCH_1997} and found that the phase-weighted stack worked best. So, all the stacks discussed here used the phase-weighted method.  Figure 2 shows an example of the envelopes of $S_{i j} (\tau)$ for the Big Skidder array for the 110 one-minute long time windows when tremor was detected in the 5 km by 5 km grid cell located 7 kilometers southwest of the array. As the energy of the P-waves is expected to be higher on the vertical component, and the energy of the S-waves to be higher on the horizontal components, we interpret this peak to correspond to the time lag between the arrival of the direct P- and S-waves.

@@

To improve the signal-to-noise ratio of the peak in the stacked cross-correlation, we divided the one-minute-long time windows into two clusters, the ones that match the stacked cross-correlations well, and the ones that do not. In order to do the clustering, we stacked the stacked cross-correlation functions, $S_{i j} (\tau)$, over all the one-minute-long time windows to obtain $S_i (\tau)$. For each 1-minute time window $j$ we cross-correlated $S_i (\tau)$ with $S_{i j} (\tau)$ to obtain $S^c_{i j} (\tau)$.  For each $i$ and $j$ we determine 4 numbers from $S^c_{i j} (\tau)$: its value at zero lag time, its maximum absolute value, the time lag at which it takes its maximum absolute value and the ratio of the maximum absolute value to the rms of the correlation function. We considered lag times from X to Y in each of these measures. Each one-minute-long time window is thus associated with eight values of quality criteria, four for each of the the East and North channels. We then classified each one-minute-long time window into two different clusters, based on the value of these criteria, using a K-means clustering algorithm (function sklearn.cluster.KMeans from the Python library SciKitLearn). The K-means procedure is as follows: We choose to have 2 clusters, then we arbitrarily choose a center for each cluster. We put each one-minute-long time window into the cluster to which it is closest (based on the values of the eight criteria). Once all one-minute-long time windows have been put in a cluster, we recompute the mean of the eight criteria for each cluster, and iterate the procedure until convergence. On average, about 35\% of the time windows fit well with the stack and are kept in the first cluster, while 65\% do not fit well with the stack and are removed. For each cluster, we then stacked the cross-correlation functions over all the one-minute-long time windows belonging to that cluster using a phase-weighted stack. We tried using more than 2 clusters, but it did not improve the final stack of the best cluster. Figure 3 shows the envelope of the stack of $S_{i j} (\tau)$ over the time windows $j$ that correspond to cluster 0 (red) and cluster 1 (blue).  The clustering has greatly improved the amplitude of the peak for cluster 1, and made the peak nearly disappear for cluster 0.

@@

In the following, we kept only the data points for which the maximum amplitude of the stack is higher than 0.05. Figures 6, 7, 8 and 9 show respectively a map and several cross-sections of the depth of the source of the tremor. Figure 10 and 11 show a map of the distance between the source of the tremor and the plate boundary from the McCrory model and a map of the distance to the plate boundary from the Preston model, alongside with the depth of the low-frequency earthquake families observed by ~\citeA{SWE_2019} and ~\citeA{CHE_2017_JGR}. There is a good agreement between the depth of the source of the tremor, the depth of the low-frequency earthquake families, and the depth of the plate boundary. On the cross-section figures, we compare the distance of the source of the tremor to the plate boundary for different locations of the cross-section. We note that the depth of the tremor matches well the depth of the low-frequency earthquakes.

@@

\begin{figure}
\noindent\includegraphics[width=\textwidth, trim={0.5cm 5cm 2cm 7.5cm},clip]{figures/section_strike_PWS_PWS_1.eps}
\caption{Cross-section showing depth of the tremor and of the low-frequency earthquake families (small filled circles) identified by ~\citeA{SWE_2019} and ~\citeA{CHE_2017_JGR} for cross section A-A' shown in Figure 6. The black line corresponds to the plate boundary profile of McCrory (2006) along the southern most black line shown on Figure 6, and the dashed line to the plate boundary profile of Preston (2003). The color bar shows the distance of the tremor and the low-frequency earthquakes along the strike. Only tremor and low-frequency earthquakes less 10km away from the profile line are shown. Tremor depths are from arrays Burnt Hill (squares), Big Skidder (stars), Danz Ranch (inverted triangles),  Gold Creek (large circles), Port Angeles (diamonds), and Three Bumps (triangles).}
\label{pngfiguresample}
\end{figure}

\begin{figure}
\noindent\includegraphics[width=\textwidth, trim={0.5cm 5cm 2cm 7.5cm},clip]{figures/section_strike_PWS_PWS_2.eps}
\caption{Cross-section showing the depth of the source of the tremor and of the low-frequency earthquake families for cross section B-B' shown in Figure 6. Symbols are described in caption for Figure 7.}
\label{pngfiguresample}
\end{figure}

\begin{figure}
\noindent\includegraphics[width=\textwidth, trim={0.5cm 5cm 2cm 7.5cm},clip]{figures/section_strike_PWS_PWS_3.eps}
\caption{Cross-section showing the depth of the source of the tremor and of the low-frequency earthquake families for cross section C-C' shown in Figure 6. Symbols are describe in caption for Figure 7.}
\label{pngfiguresample}
\end{figure}

\begin{figure}
\noindent\includegraphics[width=\textwidth, trim={1cm 5cm 3.5cm 4cm},clip]{figures/d_to_pb_PWS_PWS_P.eps}
\caption{Map of the distance between the plate boundary (from the McCrory model) of the tremor and of the low-frequency earthquake families (filled circles) identified by ~\citeA{SWE_2019} and ~\citeA{CHE_2017_JGR}. Tremor located below the plate boundary is blue, and above is red. Tremor depths are from arrays Burnt Hill (squares), Big Skidder (stars), Danz Ranch (inverted triangles),  Gold Creek (large circles), Port Angeles (diamonds), and Three Bumps (triangles).}
\label{pngfiguresample}
\end{figure}

@@

>>>>

For each array, and every 5 km by 5 km grid cell located within 25 km, we analyze the one-minute long time windows corresponding to all the tremor epicenters located within the grid cell. For each three-component seismic station and each channel, we detrended the data, tapered the first and last 5 seconds with a Hann window, removed the instrument response, bandpass filtered between 2 and 8 Hz, and resampled the data to 20 Hz. All these preprocessing operations were done with the Python package obspy. For each seismic station and each one-minute-long time window, we cross correlated the vertical component with the East-West component and with the North-South component. Then, we stacked the cross correlation functions over all the seismic stations of the array. We call this stack $S_{i j} (\tau)$ for time lag $\tau$, the $i$-th channel ($i$=1 or 2 for the North or East channel) and the $j$-th one-minute time window for a given epicenter grid/array pair. At each step of this method we experimented with a linear stack, a $n$th-root stack, and a phase-weighted stack ~\cite{SCH_1997} and found that the phase-weighted stack worked best. So, all the stacks discussed here used the phase-weighted method.  Figure 2 shows an example of the envelopes of $S_{i j} (\tau)$ for the Big Skidder array for the 110 one-minute long time windows when tremor was detected in the 5 km by 5 km grid cell located 7 kilometers southwest of the array. As the energy of the P-waves is expected to be higher on the vertical component, and the energy of the S-waves to be higher on the horizontal components, we interpret this peak to correspond to the time lag between the arrival of the direct P- and S-waves.

@@

To improve the signal-to-noise ratio of the peak in the stacked cross-correlation, we divided the one-minute-long time windows into two clusters, the ones that match the stacked cross-correlations well, and the ones that do not. In order to do the clustering, we stacked the stacked cross-correlation functions, $S_{i j} (\tau)$, over all the one-minute-long time windows to obtain $S_i (\tau)$. For each one-minute time window $j$ we cross-correlated $S_i (\tau)$ with $S_{i j} (\tau)$ to obtain $S^c_{i j} (\tau)$.  For each $i$ and $j$ we determine 3 numbers from $S^c_{i j} (\tau)$: its value at zero lag time, its maximum absolute value and the time lag at which it takes its maximum absolute value, and we determine one number from the $S_{i j} (\tau)$: the ratio of its maximum absolute value to its rms. We considered lag times from 12 to 14 seconds to compute the value of the rms. Each one-minute-long time window is thus associated with eight values of quality criteria, four for each of the the East and North channels. We then classified each one-minute-long time window into two different clusters, based on the value of these criteria, using a K-means clustering algorithm (function sklearn.cluster.KMeans from the Python library SciKitLearn). The K-means procedure is as follows: We choose to have 2 clusters, then we arbitrarily choose a center for each cluster. We put each one-minute-long time window into the cluster to which it is closest (based on the values of the eight criteria). Once all one-minute-long time windows have been put in a cluster, we recompute the mean of the eight criteria for each cluster, and iterate the procedure until convergence. On average, about 35\% of the time windows fit well with the stack and are kept in the first cluster, while 65\% do not fit well with the stack and are removed. For each cluster, we then stacked the envelopes of the cross-correlation functions over all the one-minute-long time windows belonging to that cluster using a phase-weighted stack. We tried using more than 2 clusters, but it did not improve the final stack of the best cluster. Figure 3 shows the stack of the envelopes of the $S_{i j} (\tau)$ over the time windows $j$ that correspond to cluster 0 (red) and cluster 1 (blue).  The clustering has greatly improved the amplitude of the peak for cluster 1, and made the peak nearly disappear for cluster 0.

@@

In the following, we kept only the data points for which the maximum amplitude of the stack is higher than 0.05. Figures 6 and 7 show respectively a map and several cross-sections of the depth of the source of the tremor. Figure 8 shows a map of the distance between the source of the tremor and the plate boundary from the Preston model, alongside with the depth of the low-frequency earthquake families observed by ~\citeA{SWE_2019} and ~\citeA{CHE_2017_JGR}. There is a good agreement between the depth of the source of the tremor, the depth of the low-frequency earthquake families, and the depth of the plate boundary. On the cross-section figures, we compare the distance of the source of the tremor to the plate boundary for different locations of the cross-section. We note that the depth of the tremor matches well the depth of the low-frequency earthquakes.

@@

\begin{figure}
\noindent\includegraphics[width=5cm, trim={4.5cm 0.5cm 6cm 0cm},clip, angle=270]{figures/section_strike_PWS_PWS.eps}
\caption{Cross-sections showing depth of the tremor and of the low-frequency earthquake families (small filled circles) identified by ~\citeA{SWE_2019} and ~\citeA{CHE_2017_JGR} for cross sections A-A' , B-B', and C-C' shown in Figure 6. The black line corresponds to the plate boundary profile of McCrory (2006) along the southern most black line shown on Figure 6, and the dashed line to the plate boundary profile of Preston (2003). The color bar shows the distance of the tremor and the low-frequency earthquakes along the strike. Only tremor and low-frequency earthquakes less 10km away from the profile line are shown. Tremor depths are from arrays Burnt Hill (squares), Big Skidder (stars), Danz Ranch (inverted triangles),  Gold Creek (large circles), Port Angeles (diamonds), and Three Bumps (triangles).}
\label{pngfiguresample}
\end{figure}

@@

\begin{table}
\caption{Summary of distances to plate boundary}
\centering
\begin{tabular}{l c c c}
\hline
 & Median & MAD & Q \\
\hline
Tremor depth - Preston depth & - 0.207 km & 2.623 km & 1.313 km \\
Tremor depth - McCrory depth & - 0.850 km & 2.429 km & 1.390 km \\
LFE depth - Preston depth & - 0.189 km & 1.312 km & 0.871 km \\
LFE depth - McCrory depth & - 0.495 km & 1.805 km & 0.988 km \\
\hline
\end{tabular}
\end{table}
